---
title: "Data 607 Project 2"
author: "Jayden Jiang"
date: "2025-03-10"
output: html_document
---

# Library
```{r}
library(tidyr)
library(dplyr)
```

# Load Dataset #1
```{r}
dataset1 <- read.csv("https://raw.githubusercontent.com/JaydeeJan/Data-607-Project-2/refs/heads/main/products%20across%20different%20regions.csv")

print(dataset1)
```

# Tidy and Transform the Data
```{r}
# converting the data from wide format to long format for easier analysis
long_data1 <- dataset1 %>%
  pivot_longer(cols = starts_with("Jan") :starts_with("Jun"),
               names_to = "Month",
               values_to = "Sales")

head(long_data1)
```

# Analysis
```{r}
# calculating total sales per product and per region
total_sales_product <- long_data1 %>%
  group_by(Product.Name) %>%
  summarise(Total_Sales = sum(Sales))

total_sales_region <- long_data1 %>%
  group_by(Region) %>%
  summarise(Total_Sales = sum(Sales))

print(total_sales_product)

print(total_sales_region)
```

# Conclusion
The analysis shows the total sales for each product and region. Product B has the highest total sales, and the East region has the highest total sales across all products.

# Load Dataset #2
```{r}
dataset2 <- read.csv("https://raw.githubusercontent.com/JaydeeJan/Data-607-Project-2/refs/heads/main/rolling_stone.csv")

head(dataset2)
```
# Tidy and Transform the Data
```{r}
dataset2 <- dataset2 %>%
  select(album, rank_2003, rank_2012, rank_2020)

long_data2 <- dataset2 %>%
  pivot_longer(
    cols = starts_with("rank_"),
    names_to = "Rank_Year",
    values_to = "Rank")

head(long_data2)
```
# Analysis
```{r}
# Calculate the average ranking for each album
average_ranking <- long_data2 %>%
  group_by(album) %>%
  summarise(average_rank = mean(Rank, na.rm = TRUE)) %>%
  arrange(average_rank)

head(average_ranking, 10)
```
# Conclusion
The analysis shows the top 10 albums for average ranking over 2003, 2012, and 2020. 

# Load Dataset #3
```{r}
dataset3 <- read.csv("https://raw.githubusercontent.com/JaydeeJan/Data-607-Project-2/refs/heads/main/Sales%20Data.csv")

head(dataset3)
```

# Tidy and Transform the Data
```{r}
long_data3 <- dataset3 %>%
  pivot_longer(cols = January:November, names_to = "Month", values_to = "Sales") %>%
  mutate(Sales = as.numeric(gsub("[\\$,]", "", Sales)))

head(long_data3)
```
# Analysis
```{r}
# Calculate total sales per store
total_sales_per_store <- long_data3 %>%
  group_by(Store) %>%
  summarise(Total_Sales = sum(Sales)) %>%
  arrange(desc(Total_Sales))

print(total_sales_per_store)
```
# Conclusion
The analysis shows the 5 stores in US market has the highest total sales amount comparing to other regions. 
